SubGenome Analysis

### 2017-08-04 R/SubGenome Analysis.Rmd
Updated plot function
*Not in a complete state yet

  #v1 vs. v1
  synMapOutputFile: "6807_8082.CDS-CDS.lastz.tdd10.cs0.filtered.dag.all.go_D20_g10_A5.aligncoords.gcoords.ks"
  inputDataFile: "../Data/sorghum_v1_vs_maize_v1.tab"
  outputFile: "../Output/recoveredSubGenomes_sorghum_v1_vs_maize_v1.tab"
  log10_ks_cutoff: 0
  #-0.575

  #v3.1 vs v4+rejected
  #synMapOutputFile: ""
  #outputFile: "../Output/recoveredSubGenomes_sorghum_v3.1_vs_maize_v4+rejected.tab"
  #inputDataFile: "../Data/sorghum_v3.1_vs_maize_v4+rejected.tab"
  #log10_ks_cutoff: -0.575
---

@@

We need to select a cutoff value for ks (synonymous mutation rate) to differentiate between the most recent alpha duplication event and the previous beta duplication. If the vertical line separates the first two peaks sufficently, it is probably a good enough value.  If not, we need to pick a new cutoff value for ks.

```{r pickKSCutoff, echo=FALSE}
qplot(log10(syntelogs.complete$ks), geom="histogram", binwidth=.01)  + geom_vline(mapping = NULL, data = NULL, xintercept = params$log10_ks_cutoff, na.rm = FALSE, show.legend = NA)
```

@@

```{r reportFindSubgenomes, echo=FALSE, results="asis"}
## Display chromosomes with significant syntenic genes between org1 and org2
homeologs.sorted %>%
  select(org_chr1, org_chr2, chromosomeGeneCount) %>%
  spread(org_chr2, chromosomeGeneCount) -> homeologs.spread

@@

```{r writeSubgenomesToFile, echo=FALSE}
## Now recover which genes are in these sets
write.table(subgenome, params$outputFile, sep="\t")
```

We also want to know how we did compared to the paper itself, to see if this method is faithfully reproducing their results.
```{r compareResultsToPaper, echo=FALSE}
sd01 <- read_excel("../Data/sd01.xls")
```

>>>>

  #v1 vs. v1
  # synMapOutputFile: "6807_8082.CDS-CDS.lastz.tdd10.cs0.filtered.dag.all.go_D20_g10_A5.aligncoords.gcoords.ks"
  # inputDataFile: "../Data/sorghum_v1_vs_maize_v1.tab"
  # outputFile: "../Output/recoveredSubGenomes_sorghum_v1_vs_maize_v1.tab"
  # log10_ks_cutoff: 0

  #v3.1 vs v4+rejected
  synMapOutputFile: "31607_34889.CDS-CDS.lastz.tdd10.cs0.filtered.dag.all.go_D20_g10_A5.aligncoords.gcoords"
  inputDataFile: "../Data/sorghum_v3.1_vs_maize_v4+rejected.tab"
  outputFile: "../Output/recoveredSubGenomes_sorghum_v3.1_vs_maize_v4+rejected.tab"
  log10_ks_cutoff: 0
  #-0.575
---

@@

We need to select a cutoff value for ks (synonymous mutation rate) to differentiate between the most recent alpha duplication event and the previous beta duplication. If the vertical line separates the first two peaks sufficently, it is probably a good enough value.  If not, we need to pick a new cutoff value for ks. Note that looking at ks values by gene produces no signal.  Looking at median ks values by syntenic block (kindof) shows two peaks, first for orthologs and second for homeologs from "pregrass tetraploidy".

```{r pickKSCutoff, echo=FALSE}
qplot(
  syntelogs.mutated$ks[syntelogs.mutated$ks < 2.5], 
  geom="histogram", binwidth=.05, na.rm=TRUE
  ) + geom_vline(
    mapping = NULL, 
    data = NULL, 
    xintercept = 10^params$log10_ks_cutoff, 
    na.rm = FALSE, 
    show.legend = NA
  ) + labs(
    title="Frequency of KS values by syntenic block (only KS values < 2.5)", 
    x="KS value of syntenic block", 
    y="Count (# of genes)"
  )

qplot(syntelogs.mutated$median_ks[syntelogs.mutated$median_ks < 2.5], geom="histogram", binwidth=.05, na.rm=TRUE)  + geom_vline(mapping = NULL, data = NULL, xintercept = 10^params$log10_ks_cutoff, na.rm = FALSE, show.legend = NA) + labs(title="Frequency of median KS values by syntenic block (only median KS values < 2.5)", x="Median KS value of syntenic block", y="Count (# of genes)")

qplot(log10(syntelogs.raw$ks), geom="histogram", binwidth=.01, na.rm=TRUE)  + geom_vline(mapping = NULL, data = NULL, xintercept = params$log10_ks_cutoff, na.rm = FALSE, show.legend = NA) + labs(title="Frequency of log10 KS values", x="log10(KS) value of genes", y="Count (# of genes)")
```

@@

```{r reportAncestralChromosomes, echo=FALSE, results="asis"}
## Display chromosomes with homeologous genes between org1 and org2
homeologs.chromosome %>%
  select(org_chr1, org_chr2, chromosomeGeneCount) %>%
  spread(org_chr2, chromosomeGeneCount) -> homeologs.spread

@@

```{r writeSubgenomesToFile, echo=FALSE}
## Now recover which genes are in these sets
#write.table(subgenome, params$outputFile, sep="\t")
```

We also want to know how we did compared to the paper itself, to see if this method is faithfully reproducing their results.
```{r compareResultsToPaper, echo=FALSE}
#sd01 <- read_excel("../Data/sd01.xls")
```
